---
title: "ndi_analysis"
author: "Harris Policy Lab"
date: "2/16/2021"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# List of packages to install
packages <- c("tidyverse", "janitor", "reticulate", "countrycode", "lubridate")

# To install the packages, uncomment the line below and run it
#install.packages(packages)

library(tidyverse)
library(janitor)
library(reticulate)
library(countrycode)
library(lubridate)

```

```{r importing_and_merging}

# COVID-19 data from Our World in Data
owid_covid_data <- read_csv("data/owid-covid-data.csv")
df_covid_owid <- owid_covid_data %>%
  clean_names() %>%
  mutate(
    country_standard = countryname(location),
    date = mdy(date),
    year = year(date)) %>%
  select(country_standard, year, everything())

# Economist Intelligence Unit
EIU_democracy_index_clean <- read_csv("data/EIU_democracy_index_clean.csv")
df_eiu <- EIU_democracy_index_clean %>%
  clean_names() %>%
  mutate(country_standard = countryname(country)) %>%
  # Adding "OGP" prefix to column names
  rename_with( 
  ~ paste("eiu", .x, sep = "_"), 
  overall_score:date) %>%
  select(country_standard, year, eiu_overall_score:eiu_regime_type)

# Data from Open Government Partnership
OGP_data <- read_excel("data/OGP_data.xlsx", 
    sheet = "Data")
df_ogp <- OGP_data %>%
  clean_names() %>%
  mutate(country_standard = countrycode(
    country_code, 
    origin = "iso3c", 
    destination = "country.name")) %>%
  # Adding "OGP" prefix to column names
  rename_with( 
  ~ paste("ogp", .x, sep = "_"), 
  date_joined:fiscal_openness_oversight) %>%
  select(country_standard, everything())

# Merging datasets
df_all <- df_covid_owid %>%
  full_join(df_eiu, by = c("country_standard", "year")) %>%
  full_join(df_ogp, by = "country_standard") %>%  
  select(
    country_standard, 
    year, 
    date,
    eiu_overall_score:eiu_regime_type,
    ogp_participating:ogp_fiscal_openness_oversight,
    everything()) %>%
  mutate(post_covid = 
           ifelse(year >= 2020,
                  TRUE,
                  FALSE))

# Exporting dataset
write_csv(df_all, "data/merged_data.csv")
```

Note that the following country codes don't appear to be iso3c and don't match to a country name: ARBA, BRSP, CAON, CONR, ESMD, ESPV, FRPA, GBSC, GETB, GHST, IDBO, KEEM, KRSE, MXJL, NGKD, PELL, PHSC, ROIS, TZKG, USAU

You can do python code, like in example chunk below.

```{python}
print("hello world")
```

## Resources:

-   

## Citations:

This project uses the countrycode R package

Arel-Bundock, Vincent, Nils Enevoldsen, and CJ Yetman, (2018). countrycode: An R package to convert country names and country codes. Journal of Open Source Software, 3(28), 848, <https://doi.org/10.21105/joss.00848>
